Smart Drill-Down: Interactive Summarization of Relational Tables

### 2014-11-24 TableSummarization/Paper/exp.tex
Addressed Aditya's comments. Changes in experiments.

Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit into the page. We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.

@@

\subsubsection{Weighting functions}
Our system can display optimal rule summaries using any monotonic weighting function. By default, we assign a rule weight equal to it's size. In this section, we try out a couple of other weighting functions.

@@

\subsubsection{Effects of $m_w$}
Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$. The algorithm is guaranteed to find the best marginal rule as long as it's weight is $\leq m_w$, but runs faster for smaller values of $m_w$. In this section, we study the effect of varying $m_w$ on the speed of our algorithm. Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.

@@

@@

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
\vspace{-13pt}
\end{figure}

@@

The computation time increases approximately linearly with parameter $minSS$. This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
% TODO redo the census bits weighting experiments with proper m_w (forgot to set it correctly earlier)

% TODO, find for census dataset.
In addition, we measure the number of incorrect rules per iteration. If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule. We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}. This number is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function. Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
  \caption{Running time for different values of parameter $minSS$ \label{fig:minSS_speed}}
\vspace{-13pt}
\end{figure}

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_error_percent.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
  \caption{Error in Count for different values of parameter $minSS$ \label{fig:minSS_error_percent}}
\vspace{-13pt}
\end{figure}

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_error_rule.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-10pt}
\caption{Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_error_rule}}
\vspace{-13pt}
\end{figure}

>>>>

Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit into the page. We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function. Memory capacity $M$ for the SampleHandler is set to $50000$ tuples, and $minSS$ to $5000$.

@@

\subsubsection{Weighting functions}
Our system can display optimal rule summaries using any monotonic weighting function. By default, we assign a rule weight equal to its size. In this section, we try out a couple of other weighting functions.

@@

\subsubsection{Effects of $m_w$}
Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$. The algorithm is guaranteed to find the best marginal rule as long as its weight is $\leq m_w$, but runs faster for smaller values of $m_w$. In this section, we study the effect of varying $m_w$ on the speed of our algorithm. Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.

@@

For the Census dataset, the running time is dominated by time spent in making a pass through the $2.5$ million tuples to create the first sample. The response time for the next user click should be quite small, as the sample created for the first expansion can usually be re-used for the next rule expansion.

@@

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/mw_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-5pt}
  \caption{Running time for different values of parameter $m_w$ \label{fig:mw_speed}}
\vspace{-20pt}
\end{figure}

@@

The computation time increases approximately linearly with parameter $minSS$. This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample. The running time at very small values of $minSS$ is simply the time taken to read the table and create the sample. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.

In addition, we measure the number of incorrect rules per iteration. If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule. We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}. This number for the Marketing dataset is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function. For the Census dataset, it is around $1$ for $minSS \leq 1000$ and falls to about $0.3$ for larger values of $minSS$. Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_speed.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-5pt}
  \caption{Running time for different values of parameter $minSS$ \label{fig:minSS_speed}}
\vspace{-20pt}
\end{figure}

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_error_percent.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-5pt}
  \caption{Error in Count for different values of parameter $minSS$ \label{fig:minSS_error_percent}}
\vspace{-20pt}
\end{figure}

\begin{figure}
\hspace{-20pt}
  \includegraphics[height=2in]{graphs/minSS_error_rule.pdf}%change height to 2 inch for actual paper, 5 inch for single column
\vspace{-5pt}
\caption{Average number of incorrect rules for different values of parameter $minSS$ \label{fig:minSS_error_rule}}
\vspace{-20pt}
\end{figure}

### 2014-11-24 TableSummarization/Paper/formal.tex
Addressed Aditya's comments. Changes in experiments.

Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score.

>>>>

Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules of given size $k$ that tries to minimize the length of this description, is equivalent to finding the rule-set that maximizes the total score.

### 2014-11-24 TableSummarization/Paper/related.tex
Addressed Aditya's comments. Changes in experiments.

>>>>

We use sampling to find approximate estimates of rule counts. Various other database systems~\cite{Acharya:1999:AAQ:304182.304581, Agarwal:2013:BQB:2465351.2465355} use samples to find approximate results to SQL aggregation queries. These systems create samples in advance and only update them when the database changes. In contrast, we keep updating our samples on the fly, as the user interacts with our system.
